Exercise 3: Software Architecture in Practice Case: Beer Web Store

### 2008-03-17 trunk/SAiP/Module1/H3/doc/productlines.tex
Adding SOA vs. Product-line

Defining the architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money. The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the architecture and very little (if any) redesign has to be done. As more and more products are created based on the architecture their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.

Letting the architecture evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product. When the second product is to be created more of the architecture is defined based on the parts of the first product, which may be used in the second. Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire architectur was defined in advance. As more products come along the amount of reuse grow, and would eventually reach the same as the one where everything was defined in advance.

If the architecture is used to define a Product Line it becomes even more pronounced, and the above two approaches are defined as Heavyweight and Lightweight, respectively. A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in Figure \ref{fig:product-line}.

@@

\fixme{Answer question 5.3}

>>>>

Defining the product line architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money. The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the product line and very little (if any) redesign has to be done. As more and more products are created based on the product line their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.

Letting the product line evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product. When the second product is to be created more of the product line is defined based on the parts of the first product, which may be used in the second. Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire product line was defined in advance. As more products come along the amount of reuse grow, and would eventually reach the same as the one where everything was defined in advance.

The above two approaches are defined as Heavyweight and Lightweight, respectively. A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in Figure \ref{fig:product-line}.

@@

As mentioned in Section \ref{sec:qa_q4} part of SOA is about generalization, just like product lines are. Taking the example given above, having the architecture support mobile phones (i.e. WAB) means that the server must be able to supply a scaled down result of a search as well as the fully blown. The server must thereby handle different types of clients.

The requirement that other product's such as other stamps or soda-pop should be suppoprted, means that the server's domain model and database must support general elements as opposed to just beer elements.

This generalization fits well with SOA, as it is not that clever to make a "Beer purchasing Service", but far better to create a "Purchasing service", which operate on elements. This will allow any company, who has products they wish to sell, to use this service. This is the same generalization requirement that apply to the Product line architecture, where the architecture must handle the purchase of an arbitrary product, be it stamps, soda-pop or beer. It is possible that the architecture requires that it is a quantifiable product, yet this is not an unlikely requirement for a service either.

It may now seem like SOA is the answer to creatign a product line, yet this is not the case. It is very possible to create very specific services, like the "Beer look-up service", and this may be just fine for a Beer Web Store based on SOA, yet this is not a product line. The same architectual consideration must be done when creating the elemenets of a product line and define their interrelations, must also be done when defining the services and their interrelations. The only "structure" that SOA imposes by definition is the loose coupling between the user and the service.
